########### Unhash to install Slidify ##############
## require(devtools)                              ##
## install_github("slidify", "ramnathv")          ##
## install_github("slidifyLibraries", "ramnathv") ##
## library(slidify)                               ##
##                                                ##
## Packages within Iris example:                  ##
## install.packages("caret")                      ##
## install.packages("rattle")                     ##
## library(caret); library(rattle)                ##
####################################################


---
title       : RMarkdown and Knitr 
subtitle    : UC Davis ABG 250 - Mathematical Modeling in Biological Systems
author      : Courtney D. Shelley
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : mathjax       # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---


## What is RMarkdown?

- RMarkdown is a tool to integrate writing blocks with "live" R code.

- R code is evaluated as part of the processing of the markdown document.

- Data, code, and processing can flow in a single document. 

- Reproducible research through literate statistical programming is the goal:         
    - An article which is a stream of **text** and **code** chunks, **weaved** to produce a human-readable document and **tangled** to produce a machine-readable document. 
    - R code is evaluated when the document is processed and the results are inserted into the document, demonstrating functional code and reproducible results.

- See http://rmarkdown.rstudio.com/ 

--- .class #id 



## Creating an RMarkdown Document

- If you haven't done so already, install **RStudio**

  - Go to http://www.rstudio.com/products/rstudio/download/ and select the appropriate download for your system.

- Select File $\rightarrow$ New File $\rightarrow$ R Markdown from the dropdown menu.

- Give your document a title and author.

- You have three output format options:

  - HTML produces a web-readable document, which will open in any browser (hint: great for emailing but don't use this if you plan on printing your document)
  - PDF, uneditable to receivers and maintains its formatting on any platform. 
  - Word documents allow the receiver to edit and comment on your document.

--- .class #id



## RMarkdown Document Basics 

- The front-of-document chunk within the `---` is a bit of css code that formats your document.  With a bit of research, you can customize this.  See http://rmarkdown.rstudio.com/html_document_format.html.

- Text chunks are separated from code chunks.  Text chunks can include mathematical formulas, web links, anything you'd normally include in a text editor like Word.

- Code chunks are separated within `'''{r} [code chunk] '''`.  By default, code will echo back as output before the results of the code, but this can be modified with `echo = FALSE`.  
  - See http://rmarkdown.rstudio.com/authoring_rcodechunks.html.  
  - Hint: It's a good idea to name your chunks - `'''{r nameHere}`

--- .class #id



## Text Chunk Options

```
Emphasis:     *italic* or _italic_ 
              **bold** or __bold__
   
Headers:      # Header 1    
              ## Header 2   
              ### Header 3  

Unordered lists:    - Item 1
                    - Item 2
                    - Item 3

Ordered lists:      1. Item 1
                    2. Item 2
                    3. Item 3
```

--- .class #id




## A Simple Example

```
## Classification With Iris Data Set

We will use R.A. Fisher's classic iris data set to generate a classification tree. 
```

```{r eval=FALSE}
# ```{r loadData, message = FALSE}
#Load data and required packages
data(iris)
library(caret); library(rattle)
nrow <- nrow(iris); ncol <- ncol(iris)  #inline code
iris[1:4,]
# ```
```  

```
The data consists of `'r nrow'` rows and `'r ncol'` columns, with no missing values.  
```

--- .class #id



## Classification Analysis with Iris Data Set

We will use R.A. Fisher's classic iris data set to generate a classification tree. 

```{r loadData, message = FALSE}
#Load data and required packages
data(iris)
library(caret); library(rattle)
nrow <- nrow(iris); ncol <- ncol(iris)  # inline code
iris[1:4,]
```  

The data consists of `r nrow` rows and `r ncol` columns, with no missing values.  

--- .class #id

```
### Analysis

Analysis consisted of fitting a predictive model of iris species based on 
petal length/width and sepal length/width. 
```

```{r eval=FALSE}
#```{r analysis, message = FALSE}
modFit <- train(Species ~., method = "rpart", data=iris) #Fit model
print(modFit$finalModel)   #Summarize model
#```
```

```
The final analysis is presented as a decison tree.
```
```{r, eval=FALSE}
#```{r tree, fig.width = 5, fig.height = 4}

fancyRpartPlot(modFit$finalModel) #Plot decision tree
#```
```

--- .class #id

### Analysis

Analysis consisted of fitting a predictive model of iris species based on petal length/width and sepal length/width. The final model was then plotted as a decision tree.

```{r analysis, message=FALSE}
modFit <- train(Species ~., method = "rpart", data=iris) #Fit model
print(modFit$finalModel)   #Summarize model
```

--- .class #id


The final model was then plotted as a decision tree.

```{r tree, fig.width = 12, fig.height=6}
fancyRpartPlot(modFit$finalModel) #Plot decision tree
```

Full document link: http://rpubs.com/BlackWidoww/52864

--- .class #id

## More Cool Stuff With Markdown
 
- Pandoc Markdown allows for tables, images, citations, etc.  See http://rmarkdown.rstudio.com/authoring_pandoc_markdown.html#math

- Math:
  - Contain mathematical notation between two `$` characters inline with text (i.e. - within a paragraph) or as its own paragraph between `$$`. 
  - Uses LaTeX (pronounced Lay-teck) 
  - http://www.artofproblemsolving.com/Wiki/index.php/LaTeX:Symbols
  - `f(x) = $\frac{\lambda x}{1+(ax)^b}$` $\rightarrow$ $f(x)\:=\:\frac{\lambda x}{1\:+\:(ax)^b}$
  
- Links:
  - Enclose a url between pointy brackets, `<url>`, for an automatic link.
  - Inline links: `[my link](/url)` 
  - Reference links: Here is `[my link][LINK]`    `[LINK]: /url/goes/here`
    

--- .class #id

## Knitr

Finish your document:
1. Save your document by clicking the floppy disk icon or File $\rightarrow$  Save. Your document will have the file extension `.Rmd`. 

2. Click `Knit HTML` to create two output documents:

  - A `.html` document, which will open in your favorite web browser.
  - A `.md` document, which contains all the document formatting. 
  
## RPubs  
You can publish an RMarkdown piece to a publicly accessible website using RPubs by clicking `Publish` on your finished HTML.   



Built in [Slidify](http://ramnathv.github.io/slidify/index.html). Hosted on [RPubs](http://rpubs.com/BlackWidoww). Created by [@_BlackWidoww](https://github.com/BlackWidoww)
  
