---
title: "Classification With Iris Data Set"
author: "Courtney D. Shelley"
date: "February 7, 2015"
output: html_document
---


We will use R.A. Fisher's classic iris data set to generate a classification tree.  This data set is freely available from the **UCI Machine Learing Repository**, https://archive.ics.uci.edu/ml/datasets/Iris.  

The data set includes:

1. Sepal length in cm
2. Sepal width in cm
3. Petal length in cm
4. Petal width in cm
5.  class:

    -_Iris setosa_
  
    -_Iris versicolor_
  
    -_Iris virginica_
  
### Data Exploration

```{r loadData}
#Load data

data(iris)
nrow <- nrow(iris); ncol <- ncol(iris)  #inline code
iris[1:4,]  # View first four rowss of data set
```  


### Analysis

Analysis consisted of fitting a predictive model of iris species based on 
petal length/width and sepal length/width using the `caret` package. 

```{r modFit, message=FALSE}

#Load required package:
library(caret)
modFit <- train(Species ~., method = "rpart", data=iris) #Fit model
print(modFit$finalModel)   #Summarize model
```



The final analysis is presented as a decison tree, using the `rattle` package.

```{r plotTree, fig.width=6, fig.height=4, message=FALSE}
library(rattle)
fancyRpartPlot(modFit$finalModel) # Plot decision tree
```

## Discussion

_I. setosa_ can be successfully isolated from other species by petal length.  _I. versicolor_ and _I. virgicina_ are somewhat more difficult to separate based on petal width.  



This document is hosted freely on RPubs:   http://rpubs.com/BlackWidoww/58755

.Rmd file and .pdf versions can be found here:  https://github.com/BlackWidoww/ABG250
